Evaluation of Design Trade-offs for Adders\\ in Approximate Datapath

### 2014-11-23 WAP2015/wap2015.tex
submitted version

\usepackage{algorithm}
\usepackage{array}
\usepackage{algpseudocode}
\usepackage{indentfirst}
\usepackage[pdftex]{graphicx}
\usepackage[tight,footnotesize]{subfigure}

@@

\author{\IEEEauthorblockN{Kan Shi and George A. Constantinides}
\IEEEauthorblockA{Department of Electrical and Electronic Engineering\\
Imperial College London\\
London, UK\\
Email: \{k.shi11, g.contantinides\}@imperial.ac.uk}
\vspace{2ex}
%\and
%\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
%\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
%line 2: name of organization, acronyms acceptable\\
%line 3: City, Country\\
%line 4: Email: name@xyz.com}
}

@@

\begin{abstract}
Releasing the stringent accuracy requirement would potentially offer greater freedom to create a design with better performance or energy efficiency. In this paper, we evaluate the design trade-offs for adders, which are key building blocks for many applications. We demonstrate the optimum design metric for adders, under the consideration of various design constraints, such as accuracy, operating frequency, silicon area, different types of adder structure and even alternative form of computer arithmetic for adder implementation. Two design scenarios are compared: one is the conventional design scenario where timing closure is met by adjusting operand word-lengths. The other is an overclocking scenario where timing violations are allowed to occur. We show that applying the overclocking approach to ripple carry adders can be more beneficial than using fast adders to achieve similar latency, because the worst cases only happen with very small probabilities. We also show that using the conventional approach on adders with new form of computer arithmetic is optimal for a wide range of design constraints.

@@

\section{Introduction}
Circuit performance has been increased tremendously over the past decades. However, in recent years people have observed an end of performance improvement~\cite{DarkSilicon2011} due to the scaling of CMOS technology to the nanometer regime. In order to boost circuit performance, two standard design techniques are commonly adopted: one is to heavily pipeline the datapath, the other is to reduce the design precision. For the first technique, pipelining can be used to increase operating frequency, whereas the overall computational latency will not be affected. Therefore the application of this technique is limited in many embedded applications, which are often designed with strict latency specifications. The second technique is common especially with FPGA technology, which embodies the flexibility to implement customized variable representation and is able to optimize the operand word-length across the entire datapath to meet a given performance and area constraint. However, the precision loss will inevitably introduce quantization errors into the design.

Unfortunately, in both cases circuits are designed for the worst cases. In order to avoid potential timing violations, timing analysis tools tend to provide a more conservative path delay by adding guard bands and timing margins. However, continuing with this approach will become increasing expensive and difficult as technology shrinks.

To tackle this problem, a large volume of research activities were focused on the design methodology of relaxing the $100\%$ accuracy requirements and design constraints, while evaluating the corresponding design benefits and costs. Studies have shown that computing approximately provides extra opportunities to design circuits with better performance and less energy consumption. A brief discussion of the work in this area is given in Section~\ref{sec:Background}.

In this paper, we provide detailed evaluations of design trade-offs for adders, which are key arithmetic primitives for a broad range of applications. In order to provide the optimum design choice for adders, we examine several design constrains and options, such as accuracy, operating frequency, silicon area, different types of adder structure, as well as different forms of computer arithmetic for adder implementation. In comparison to the conventional design scenario that timing closure is achieved by truncating the word-length of operands, we take a radical step forward by adopting an alternative design scenario, which is named ``overclockign scenario''. In this scenario, circuits are operated far beyond the conservation region with full precision, while timing violation can be tolerated. Previous work has shown that for many applications, it is preferable to follow the overclocking scenario because the worst case is triggered by certain input patterns and hence it happens rarely~\cite{SKfccm13}. In addition to the adder structures that are created based on standard binary arithmetic, such as ripple carry adder and carry select adder, we also evaluate adder with ``online arithmetic''~\cite{Ercegovac_OnlineOverview}, which employs different style of data representation to eliminate carry propagation in conventional adders. We demonstrate experimentally on FPGAs that for limited area budgets, applying the overclocking approach to ripple carry adders can be more beneficial than using fast adders to achieve similar latency. With more relaxed area constraints, the online adder outperforms in terms of accuracy or performance. Potentially our study will provide circuit designers with guidance and extra options when balancing between various design trade-offs.

In summary, the main contributions of this paper are:
\begin{enumerate}
  \item Detailed evaluation of three adder structures: ripple carry adder, carry select adder and online adder; 
  \item Evaluation of design scenarios and identification of the optimum design choice under the consideration of accuracy-performance-area trade-offs.
\end{enumerate}

The rest of this paper is organized as below: we first provide a brief overview of the relevant work in this area, together with the background knowledge of online arithmetic in Section~\ref{sec:Background}. The comparison between different adder structures and design trade-offs are described in detail in Section~\ref{sec:diff_adder}. This is followed by the evaluation of optimum design choices when considering trade-offs between accuracy, performance and area in Section~\ref{sec:optimum_adder}. The conclusion and identification of possible future work is given in Section~\ref{sec:conclusion}.

\section{Background}\label{sec:Background}
\subsection{Approximate Datapath Design}
The International Technology Roadmap for Semiconductors (ITRS) has identified that additional benefits of manufacturing, test, power and timing can be achieved by releasing the requirement of absolute correctness~\cite{semiconductor2007international}. A series of studies in this area called ``Better Than Worst-Case Design'' focused on creating circuits by operating beyond the safety margins. Typically this can be achieved by either reducing the supply voltage~\cite{NonUniformScaling} or increasing the operating frequency beyond the rated value~\cite{SKfccm13,JoshFCCM2012}. The Razor project~\cite{Razor2004} served as one of the exemplary designs, in which both supply voltage and clock frequency were scaled over the most conservative value, while the error rate was monitored by using a self-checking circuit. It was shown that over $30\%$ power saving can be obtained at the cost of around $1\%$ error rates.

Another stream of research was focus on proposing either approximate circuit structures or design tools by trading accuracy to performance and/or energy benefits. Work in this area was inspired by the fact that many applications, e.g. image processing and machine learning, were naturally error resilient. For example, code analysis tools~\cite{Truffle2012Uwash,EnerJ2011Uwash} were developed to divide the program into precise parts and approximate parts, which can then be mapped to hardware components with different speed-grades and supply voltages. As for the approximate hardware, numerous forms of architectures were designed for basic arithmetic operators such as adders [xxx] and multipliers [xxx]. For example, a under-designed multiplier unit was proposed, of which the worst case was replaced by a normal case based on straightforward Karnaugh-Map analysis~\cite{Undersigned2x2multiplier}. However, many of these techniques are designed at gate-level or transistor-level, and they cannot be directly applied onto existing hardware platforms such as FPGAs.

In this paper, we overcome the previous limitations by comparing three adder structures: ripple carry adder and carry select adder, which are widely employed in mainstream hardware platforms, together with online adder which can be also found when designing fast multipliers and embedded processors~\cite{RedundantMult_1987}. It is argued that our approach can be efficiently applied to existing designs, and we back up this hypothesis based on FPGA experiments.

@@

As an alternative form of computer arithmetic, online arithmetic has been used in numerous applications such as signal processing and control algorithms~\cite{Online_FPGADSP,Online_Control}. Online arithmetic was originally designed for digit-serial operation, as illustrated in Fig~\ref{Fig:OnlineDataFlow}. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required, where $\delta$ is called the ``online delay''. $\delta$ is normally a small constant, which is independent of the precision. For ease of discussion, for the rest of this paper, the input data is assumed to be fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}), where $j\in[-\delta,N-1]$ and $r$ denotes the radix \cite{Ercegovac_Book}.
%
\begin{figure}[tbp]
  \centering
  %\vspace{-2.5ex}
  \includegraphics[width=.42\textwidth]{./figures/OnlineArithmetic_DataFlow.eps}
  %\vspace{-1ex}
  \caption{Dataflow in digit-serial online arithmetic, in which both inputs and outputs are processed from the MSD to the LSD. $\delta$ denotes the online delay.}
  \vspace{-2ex}
  \label{Fig:OnlineDataFlow}
\end{figure}
%
\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

@@

\section{Design Trade-offs of Different Adder Structures}\label{sec:diff_adder}
\subsection{Ripple Carry Adder}
Adders serve as a key building block for arithmetic operations. In general, the ripple carry adder (RCA) is the most straightforward and widely used adder structure. As such, in our previous work we proposed probabilistic models of overclocking errors for RCA.

For an $n$-bit RCA, it is composed of $n$ serial-connected full adders (FA). Typically the maximum frequency of RCA is determined by the longest carry propagation. Under the assumption that the carry propagation delay of each FA is a constant value $\mu$, the critical path of the RCA is: $\mu_{RCA}=n\mu$. For the sampling period $T_S$, it follows that if $T_S\geqslant\mu_{RCA}$, correct result will always be sampled; otherwise if $T_S<\mu_{RCA}$, intermediate be sampled and potentially generating errors. In the previous work~\cite{SKfccm13}, the mean value of overclocking error is modeled as given in (\ref{Eq:MeanError_RCA}), where coefficient $b$ is given in (\ref{Eq:b}) and it determines the maximum length of error-free carry propagation in bits.
%
\begin{eqnarray}\label{Eq:MeanError_RCA}
    E_O=\left\{
        \begin{matrix}
            2^{-b}-2^{-n-1}, & \textrm{if $b\leq n$}\\
            0, & \textrm{otherwise}
        \end{matrix}
        \right.
\end{eqnarray}
%
\begin{eqnarray}\label{Eq:b}
    b:=\left\lceil \frac{T_S}{\mu} \right\rceil=\left\lceil \frac{1}{\mu\cdot f_S}\right\rceil
\end{eqnarray}

On the other hand, a specific timing target can be met by truncating the word-length of RCA while timing violation is not permitted. This will result in truncation error for most data.  The mean value of truncation error is also modeled as given in~(\ref{Eq:MeanTruncError_RCA})~\cite{SKfccm13}, where parameters $k$ and $n$ denote the word-length of RCA before and after truncation, respectively. 
%
\begin{eqnarray}\label{Eq:MeanTruncError_RCA}
    E_O=\left\{
        \begin{matrix}
            2^{-n}-2^{-k}, & \textrm{if $n<k$}\\
            0, & \textrm{otherwise}
        \end{matrix}
        \right.
\end{eqnarray}

In this paper we consider two design scenarios for RCA: one is the overclocking scenario where timing violation is allowed to happen while maintaining the original word-length; the other is the trad scenario by truncating RCA word-length to meet timing. For a given $f_S$, in the first design scenario we have $n=k$. In the second design scenario we have $n=b-1$ to prevent timing violation while minimizing the value of $E_O$. Therefore the comparison between the two scenarios in mean error value is given in (\ref{Eq:ErrorCompare_RCA}) by updating (\ref{Eq:MeanError_RCA}) and (\ref{Eq:MeanTruncError_RCA}) respectively. As seen in (\ref{Eq:ErrorCompare_RCA}), in theory the mean value of error in the overclocking scenario ($E_{oc}$) is two times smaller than that in the traditional scenario ($E_{trad}$).
%
\begin{eqnarray}\label{Eq:ErrorCompare_RCA}
    \frac{E_{oc}}{E_{trad}} = \frac{2^{-b}-2^{-n-1}|_{n=k}}{2^{-n}-2^{-k}|_{n=b-1}}=\frac{1}{2}
\end{eqnarray}

\subsection{Carry Select Adder}
Although smaller value of mean error can be achieved in the overclocking scenario, it costs extra area because the full precision is maintained. Instead, alternative adder structures, such as carry select adder (CSA), are originally designed to trade silicon area for low latency. In a CSA, the carry chain is divided into multiple overlapped stages, and each stage contains two RCAs and two multiplexers, as shown in Fig. xxx(a) and Fig. xxx(b). For a given input, two additions are performed simultaneously within a single stage where the carry input is zero and one separately. One of these two results is then selected according to the actual carry input. Although this structure brings performance benefits, it costs extra hardware resources compared to a standard RCA because the carry chain is duplicated. Furthermore, multiplexers are area-expensive to implement with FPGA technology.

\subsection{Online Adder}\label{subsec:online_adder}
In addition to standard binary arithmetic, alternative forms of computer arithmetic was studied to boost performance. For example, adder with online arithmetic is also designed for low latency at the cost of more silicon area. The structure of a digit-parallel online adder (OA) where all signals represented with SD numbers of digit set $\{-1,0,1\}$ is shown in Fig.~\ref{Fig:Radix2SD_adder}. The module ``3:2'' denotes a 3:2 compressor, which takes three inputs and generates two outputs, and is logically equivalent to a full adder (FA). A major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labelled in Fig.~\ref{Fig:Radix2SD_adder}, ideally the computation delay of this adder is only two FA delays for any operand word-length, at the expenses of one extra FA for each digit of operands. 
%
\begin{figure}[tbp]
  \centering
  \includegraphics[width=.45\textwidth]{./figures/SDAdder.eps}
  \caption{An $N$-digit binary digit-parallel online adder. Both inputs and outputs are represented using SD representation. ``3:2'' denotes a 3:2 compressor.}
  \label{Fig:Radix2SD_adder}
\end{figure}

\subsection{Design Trade-offs for Adders}
In this section, we initially demonstrate the benefits of CSA and OA in accuracy and performance, as well as the area overhead in comparison to RCA. As an example, the maximum word-lengths with respect to a variety of operating frequencies for RCA, CSA and OA are illustrated in Fig.~\ref{Fig:max_wl_adder}, where overclocking error is not allowed to happen and the conventional design scenario is evaluated for all structures. Also notice that in the experiment we consider the word-length of adders from 4-bit to 32-bit with a step of 4-bit. For a finer grain experiment, the trend will be similar to that in Fig.~\ref{Fig:max_wl_adder}.
%
\begin{figure}[tbp]
  \centering
  \includegraphics[width=.5\textwidth]{./figures/exp/max_wl.eps}
  \vspace{-4ex}
  \caption{The maximum word-lengths of different adder structures with respect to a variety of frequencies. The results are obtained from Xilinx ISE 14.7.}
  \label{Fig:max_wl_adder}
\end{figure}

As can be seen in Fig.~\ref{Fig:max_wl_adder}, for a relatively relaxed frequency requirement, both CSA and OA can be implemented with greater word-lengths than RCA. For CSA, this is because the stage parallelism enables a larger word-length, even though the multiplexer delay limits the word-length of each stage in the CSA in comparison to RCA. For OA, an even larger gap in word-length can be observed when compared to RCA. Additionally, unlike RCA and CSA of which the word-length drops gradually with the increment of frequency, OA maintains the full precision across a large rang of frequencies. This is expected, because the critical path delay of OA is theoretically irrelevant to the operand word-length. However, we also observe that for a large frequency requirement, the word-length of OA drops drastically, while RCA and CSA with 2 stages outperform. This is because the multiplexer delay becomes comparable to the delay of the carry chain for CSA, and it inhibits the benefits of parallelism.

We also record the corresponding area consumed by each adder structure in terms of look-up-tables (LUTs) in FPGA as depicted in Fig.~\ref{Fig:area_adder}. It can be seen that for a given frequency, CSA costs up to $4.5\times$ (4 stages) and $3.1\times$ (2 stages) area than RCA, and OA is up to $3.8\times$ larger than RCA for some frequency values. Therefore, we incorporate silicon area as another design metric besides accuracy and performance, and evaluate the optimum adder structure under various design trade-offs. In this paper, we examine both the overclocking scenario and the conventional scenario for RCA, while only considering the conventional scenario for other adder structures.
%
\begin{figure}[tbp]
  \centering
  \includegraphics[width=.5\textwidth]{./figures/exp/area_adders.eps}
  \caption{The maximum word-length of different adder structures.}
  \label{Fig:area_adder}
\end{figure}

\section{Evaluation of Optimum Adder Structure}\label{sec:optimum_adder}
In general, the evaluation could be of interest to a circuit designer in several ways. For example, suppose the algorithm designer wish the circuit to run at a given frequency within a given area budget, while achieving the minimum output error. Besides, for many applications, the circuit design would wish the circuit to operate as fast as possible with the minimum resource usage, whilst a certain error budget can be tolerated. In both cases, decisions must then be made on which adder structure achieves this minimum, and whether the circuit should be overclocked.

\subsection{Optimum Adder Design Metric with Given Frequency and Area Requirements}
For this type of applications, the available area and expected operating frequency are given at the design time. As an example slice through the design space, in Fig.~\ref{Fig:Error_Area} we record the mean relative error (MRE) with respect to a range of operating frequencies for different design scenarios, when the area budget is set to 35 LUTs and 55 LUTs separately. MRE can be calculated as given by (\ref{Eq:MRE}), where $E_{error}$ and $E_{out}$ denote the mean value of error and the mean value of correct outputs, respectively. Notice that the optimum adder design metric that achieves the minimum error at outputs is labeled. 
%
\begin{eqnarray}\label{Eq:MRE}
  MRE=\frac{E_{error}}{E_{out}}\times100\%
\end{eqnarray}

>>>>

\usepackage{algorithm}
\usepackage{array}
\usepackage{algpseudocode}
\usepackage{indentfirst}
\usepackage[pdftex]{graphicx}
\usepackage[tight,footnotesize]{subfigure}
\usepackage{multirow}

@@

\author{\IEEEauthorblockN{Kan Shi and George A. Constantinides}
\IEEEauthorblockA{Department of Electrical and Electronic Engineering\\
Imperial College London\\
London, UK\\
Email: \{k.shi11, g.contantinides\}@imperial.ac.uk}
\vspace{1ex}
%\and
%\IEEEauthorblockN{Authors Name/s per 2nd Affiliation (Author)}
%\IEEEauthorblockA{line 1 (of Affiliation): dept. name of organization\\
%line 2: name of organization, acronyms acceptable\\
%line 3: City, Country\\
%line 4: Email: name@xyz.com}
}

@@

\begin{abstract}
Releasing the stringent accuracy requirement would potentially offer greater freedom to create a design with better performance or energy efficiency. In this paper, we evaluate the design trade-offs for adders, which are key building blocks for many applications. We demonstrate the optimum design metric for adders, under the consideration of various design constraints, such as accuracy, operating frequency, silicon area, different types of adder structure and alternative form of computer arithmetic for adder implementation. Two design scenarios are compared: one is the conventional design scenario where timing closure is met by adjusting operand word-lengths. The other is an overclocking scenario where timing violations are allowed to occur. We show that applying the overclocking approach to ripple carry adders can be more beneficial than using fast adders to achieve similar latency, because the worst cases only happen with very small probabilities. We also show that using the conventional approach on adders with new form of computer arithmetic is optimal for a wide range of design constraints.

@@

\section{Introduction}
Circuit performance has increased tremendously over the past decades. However, in recent years people have observed an end of performance scaling~\cite{DarkSilicon2011} due to the improvement of CMOS technology to the nanometer regime. In order to boost circuit performance, two standard design techniques are commonly adopted: one is to heavily pipeline the datapath, the other is to reduce the design precision. For the first technique, pipelining can be used to increase operating frequency, whereas the overall computational latency will not be affected. Therefore the utilization of this technique is limited in many embedded applications, which are often designed with strict latency specifications. The second technique is common especially with FPGA technology, which embodies the flexibility to implement customized variable representation and is able to optimize the operand word-length across the entire datapath to meet a given performance and area constraint~\cite{GeorgeDT11}. However, the precision loss will inevitably introduce quantization errors into the design.

Unfortunately, in both situations circuits are designed for the worst cases. In order to avoid potential timing violations, timing analysis tools tend to provide a conservative critical path delay by adding guard bands and timing margins. However, continuing with this approach will become increasingly expensive and difficult as technology shrinks.

To tackle this problem, a large volume of research activities were focused on the design methodology of relaxing the $100\%$ accuracy requirements and design constraints, while reasoning about the corresponding benefits and costs. Studies have shown that computing approximately provides extra opportunities to design circuits with better performance and less energy consumption. A brief review of the work in this area is given in Section~\ref{sec:Background}.

In this paper, we provide detailed evaluations of design trade-offs for adders, which are key arithmetic primitives for a broad range of applications. In order to provide the optimum design choice for adders, we examine several design constrains and options, such as accuracy, operating frequency, silicon area, different types of adder structure, as well as different forms of computer arithmetic for adder implementation. In comparison to the conventional design scenario that timing closure is achieved by truncating the operand word-length, we take a radical step forward by adopting an alternative design scenario, which is named ``overclockign scenario''. In this case, circuits are operated far beyond the conservation region with full precision, while timing violation can be tolerated. Previous work has shown that for many applications, it is preferable to follow the overclocking scenario because the worst case is triggered only by certain input patterns and hence it happens rarely~\cite{SKfccm13}. In addition to the adder structures that are created based on standard binary arithmetic, such as ripple carry adder and carry select adder, we also evaluate adder with ``online arithmetic''~\cite{Ercegovac_OnlineOverview}, which employs different style of data representation to eliminate carry propagation. We demonstrate experimentally on FPGAs that for limited area budgets, applying the overclocking approach to ripple carry adders can be more beneficial than using fast adders to achieve similar latency. With more relaxed area constraints, the online adder outperforms in terms of accuracy or performance. Potentially our study will provide circuit designers with guidance and extra options when balancing between various design trade-offs.

In summary, the main contributions of this paper are:
\begin{enumerate}
  \item Detailed evaluation of three adder structures: ripple carry adder, carry select adder and online adder; 
  \item Comparison of design scenarios and demonstration of the optimum design choice under the consideration of trade-offs between accuracy, performance and area.
\end{enumerate}

The rest of this paper is organized as follows: we first provide a brief overview of the relevant work in this area, together with the background knowledge of online arithmetic in Section~\ref{sec:Background}. The comparison between different adder structures and design trade-offs are detailed in Section~\ref{sec:diff_adder}. This is followed by the evaluation of optimum design choices when considering trade-offs between accuracy, performance and area in Section~\ref{sec:optimum_adder}. The conclusion and suggestions of possible future work is given in Section~\ref{sec:conclusion}.

\section{Background}\label{sec:Background}
\subsection{Approximate Datapath Design}
The International Technology Roadmap for Semiconductors (ITRS) has identified that additional benefits of manufacturing, test, power and timing can be achieved by releasing the requirement of absolute correctness~\cite{semiconductor2007international}. This fact has inspired a series of studies to create circuits operating beyond the worst cases and the safety margins. Typically this can be achieved by either reducing the supply voltage~\cite{NonUniformScaling} or increasing the operating frequency beyond the rated value~\cite{SKfccm13,JoshFCCM2012}. The Razor project~\cite{Razor2004} served as one of the exemplary designs, in which both supply voltage and clock frequency were scaled over the most conservative value, while the error rate was monitored by using a self-checking circuit. It was shown that over $30\%$ power saving can be obtained at the cost of around $1\%$ errors.

Another stream of research was focusing on proposing either approximate circuit structures or design tools to trade accuracy for performance and/or energy benefits. Work in this area was inspired by the fact that many applications, e.g. image processing and machine learning, were naturally error resilient. For example, code analysis tools~\cite{Truffle2012Uwash,EnerJ2011Uwash} were developed to divide a program into precise parts and approximate parts, which can then be mapped to hardware components with different speed-grades and supply voltages. As for the approximate hardware, numerous forms of architectures were designed for basic arithmetic operators such as adders~\cite{ApproximateAdderTC} and multipliers~\cite{ApproximateMultDATE}. For example, a under-designed multiplier unit was proposed, of which the worst case was replaced by a normal case based on straightforward Karnaugh-Map analysis~\cite{Undersigned2x2multiplier}. However, many of these techniques are designed at gate-level or transistor-level, and they cannot be directly applied onto existing hardware platforms such as FPGAs.

In this paper, we overcome the previous limitations by comparing two standard adder structures: ripple carry adder and carry select adder, which are widely employed in mainstream hardware platforms. We also compare a type of adder with ``online arithmetic'', which can be also found when designing fast multipliers and embedded processors~\cite{RedundantMult_1987}. It is argued that our approach can be efficiently applied to existing designs, and we back up this hypothesis based on FPGA experiments.

@@

As an alternative form of computer arithmetic, online arithmetic has been used in numerous applications such as signal processing and control algorithms~\cite{Online_FPGADSP,Online_Control}. Online arithmetic was originally designed for digit-serial operation, as illustrated in Fig~\ref{Fig:OnlineDataFlow}. It can be seen that in order to generate the first output digit, $\delta$ digits of inputs are required, where $\delta$ is called the ``online delay''. $\delta$ is normally a small constant, which is independent of the precision. For ease of discussion, for the rest of this paper, the input data is assumed to be fixed point numbers in the range $(-1,1)$. Based on this premise, the online representation of $N$-digit operands and result at iteration $j$ are given by (\ref{Eq:Online_Operands}), where $j\in[-\delta,N-1]$ and $r$ denotes the radix \cite{Ercegovac_Book}.
%
\begin{figure}[tbp]
  \centering
  %\vspace{-2.5ex}
  \includegraphics[width=.42\textwidth]{./figures/OnlineArithmetic_DataFlow.eps}
  %\vspace{-1ex}
  \caption{Dataflow in digit-serial online arithmetic, in which both inputs and outputs are processed from the MSD to the LSD. $\delta$ denotes the online delay.}
  %\vspace{-2ex}
  \label{Fig:OnlineDataFlow}
\end{figure}
%
\begin{eqnarray}\label{Eq:Online_Operands}
\footnotesize
  X_{[j]}=\sum_{i=1}^{j+\delta}x_ir^{-i},~Y_{[j]}=\sum_{i=1}^{j+\delta}y_ir^{-i},~Z_{[j]}=\sum_{i=1}^{j}z_ir^{-i}
\normalsize
\end{eqnarray}

@@

\section{Design Trade-offs of Different Adder Structures}\label{sec:diff_adder}
\subsection{Ripple Carry Adder}\label{subsec:rca}
Adders serve as a key building block for arithmetic operations. In general, the ripple carry adder (RCA) is the most straightforward and widely used adder structure. As such, in our previous work we proposed probabilistic models of overclocking errors for RCA~\cite{SKfccm13}. A brief summary of the models is described below.
%
\begin{figure}[tbp]
  \centering
  \includegraphics[width=.5\textwidth]{./figures/RCA.pdf}
  \vspace{-4ex}
  \caption{An $n$-bit ripple carry adder. $\mu_i$ and $\mu_{RCA}$ denote the propagation delay of a full adder and the overall delay of RCA, respectively.}
  \vspace{-2ex}
  \label{Fig:rca}
\end{figure}

For an $n$-bit RCA, it is composed of $n$ serial-connected full adders (FA) as shown in Fig.~\ref{Fig:rca}. Typically the maximum frequency of RCA is determined by the longest carry propagation. Under the assumption that the carry propagation delay of each FA is a constant value $\mu_i=\mu$, the critical path of the RCA is: $\mu_{RCA}=n\mu$. For the sampling period $T_S$, it follows that if $T_S\geqslant\mu_{RCA}$, correct result will always be sampled. Otherwise, intermediate result will be sampled and potentially generating overclocking errors. The mean value of overclocking error is modeled~\cite{SKfccm13} as given by (\ref{Eq:MeanError_RCA}), where coefficient $b$ is given in (\ref{Eq:b}) and it determines the maximum length of error-free carry propagation.
%
\begin{eqnarray}\label{Eq:MeanError_RCA}
    E_O=\left\{
        \begin{matrix}
            2^{-b}-2^{-n-1}, & \textrm{if $b\leq n$}\\
            0, & \textrm{otherwise}
        \end{matrix}
        \right.
\end{eqnarray}
\vspace{-1.5ex}
%
\begin{eqnarray}\label{Eq:b}
    b:=\left\lceil \frac{T_S}{\mu} \right\rceil=\left\lceil \frac{1}{\mu\cdot f_S}\right\rceil
\end{eqnarray}
\vspace{-1ex}
\normalsize

In the conventional scenario, a specific timing target can be met by truncating the word-length of RCA while timing violation is not permitted. This will result in truncation error for most data.  The mean value of truncation error is also modeled~\cite{SKfccm13} as given by~(\ref{Eq:MeanTruncError_RCA}), where parameters $k$ and $n$ denote the word-length of RCA before and after truncation, respectively. 
%
\begin{eqnarray}\label{Eq:MeanTruncError_RCA}
    E_O=\left\{
        \begin{matrix}
            2^{-n}-2^{-k}, & \textrm{if $n<k$}\\
            0, & \textrm{otherwise}
        \end{matrix}
        \right.
\end{eqnarray}
\vspace{-2ex}

In this paper we consider two design scenarios for RCA: one is the overclocking scenario where timing violation is allowed to happen while maintaining the original word-length; the other is the traditional scenario by truncating RCA word-length to meet timing. For a given $f_S$, in the first design scenario we have $n=k$. In the second design scenario we have $n=b-1$ to prevent timing violation while minimizing the value of $E_O$. Therefore the comparison between the two scenarios in mean error value is given in (\ref{Eq:ErrorCompare_RCA}) by updating (\ref{Eq:MeanError_RCA}) and (\ref{Eq:MeanTruncError_RCA}) respectively. As seen in (\ref{Eq:ErrorCompare_RCA}), in theory the mean value of error in the overclocking scenario ($E_{oc}$) is constantly two times smaller than that in the traditional scenario ($E_{trad}$).
%
\begin{eqnarray}\label{Eq:ErrorCompare_RCA}
    \frac{E_{oc}}{E_{trad}} = \frac{2^{-b}-2^{-n-1}|_{n=k}}{2^{-n}-2^{-k}|_{n=b-1}}=\frac{1}{2}
\end{eqnarray}

\subsection{Carry Select Adder}
Although smaller value of mean error can be achieved in the overclocking scenario, it costs extra area because the full precision is maintained. Instead, alternative adder structures, such as carry select adder (CSA), are originally designed to trade silicon area for low latency. In a CSA, the carry chain is divided into multiple overlapped stages, and each stage contains two RCAs and two multiplexers. For a given input, two additions are performed simultaneously within a single stage where the carry input is zero and one separately. One of these two results is then selected according to the actual carry input. Although this structure brings performance benefits, it costs extra hardware resources compared to a standard RCA because the carry chain is duplicated. Furthermore, multiplexers are area-expensive to implement with FPGA technology.

\subsection{Online Adder}\label{subsec:online_adder}
In addition to standard binary arithmetic, alternative forms of computer arithmetic was studied to boost performance. For example, adder with online arithmetic is also designed for low latency at the cost of extra silicon area. The structure of a digit-parallel online adder (OA) where all signals represented with SD numbers of digit set $\{-1,0,1\}$ is shown in Fig.~\ref{Fig:Radix2SD_adder}. The module ``3:2'' denotes a 3:2 compressor, which takes three inputs and generates two outputs, and is logically equivalent to a full adder (FA). A major advantage of the redundant number system over the standard ripple-carry based arithmetic is that the propagation of carry is eliminated, resulting in a precision-independent computation time for addition. As labeled in Fig.~\ref{Fig:Radix2SD_adder}, ideally the computation delay of this adder is only two FA delays for any operand word-length, at the expenses of one extra FA for each digit of operands. 
%
\begin{figure}[tbp]
  \centering
  \includegraphics[width=.45\textwidth]{./figures/SDAdder.eps}
  \vspace{-1ex}
  \caption{An $N$-digit binary digit-parallel online adder. Both inputs and outputs are represented using SD representation. ``3:2'' denotes a 3:2 compressor.}
  \label{Fig:Radix2SD_adder}
  \vspace{-2ex}
\end{figure}

\subsection{Design Trade-offs for Adders}
In this section, we demonstrate the benefits of CSA and OA in accuracy and performance, as well as the area overhead in comparison to RCA. As an example, the maximum word-lengths with respect to a variety of operating frequencies for RCA, CSA and OA are illustrated in Fig.~\ref{Fig:max_wl_adder}. In this experiment, circuits are operated with timing closure, i.e. the conventional design scenario is evaluated for all structures. We consider the operand word-length of adders ranging from 32-bit to 4-bit.
%
\begin{figure}[tbp]
  \centering
  \includegraphics[width=.5\textwidth]{./figures/exp/max_wl.eps}
  \vspace{-4ex}
  \caption{The maximum word-lengths of different adder structures with respect to a variety of frequencies. The results are obtained from Xilinx ISE 14.7.}
  \label{Fig:max_wl_adder}
  \vspace{-2ex}
\end{figure}

As can be seen in Fig.~\ref{Fig:max_wl_adder}, for a relatively relaxed frequency requirement (e.g. 470 MHz), both CSA and OA can be implemented with greater word-lengths than RCA. For CSA, this is because the stage parallelism enables a larger word-length, even though the multiplexer delay limits the precision of each CSA stage in comparison to RCA. For OA, an even larger gap in word-length can be observed when compared to RCA. Additionally, unlike RCA and CSA of which the maximum word-length drops gradually with the increment of frequency, OA maintains full precision across a large range of frequencies. This is expected, because the critical path delay of OA is theoretically irrelevant to the operand word-length as discussed in Section~\ref{subsec:online_adder}. However, we also observe that given a large frequency requirement, the word-length of OA drops drastically. Besides, CSA cannot be implemented with larger precision than RCA either, because in this case the multiplexer delay becomes comparable to the delay of the carry chain for CSA, and it inhibits the benefits of parallelism. 
%
\begin{figure}[tbp]
  \centering
  \includegraphics[width=.5\textwidth]{./figures/exp/area_adders.eps}
  \vspace{-4ex}
  \caption{The maximum word-length of different adder structures.}
  %\vspace{-2ex}
  \label{Fig:area_adder}
\end{figure}

We also record the corresponding area consumption of each adder structure in terms of look-up-tables (LUTs) in FPGA as depicted in Fig.~\ref{Fig:area_adder}. It can be seen that the CSA with 4 stages and 2 stages costs up to $4.5\times$ and $3.1\times$ area than RCA, respectively. OA is also up to $3.8\times$ larger than RCA. Therefore, it is necessary to provide comprehensive evaluations of different adders by considering the trade-offs between accuracy, performance and area consumption. In summary, the evaluated adder structures as well as design scenarios are summarized in Table~\ref{Table:comparison}.

\begin{table}[tbp]
\renewcommand{\arraystretch}{1.1}
\caption{Adder Structures and Design Scenarios Considered in this Paper.}
\vspace{-1.5ex}
\label{Table:comparison}
\begin{center}
\small
\begin{tabular}{c|cc}
\hline
\hline
\multirow{2}*{\textbf{Adders}} & \multicolumn{2}{c}{\textbf{Design Scenarios}} \\
& Overclocking & Traditional \\
\hline
Ripple Carry Adder & $\surd$ & $\surd$ \\
Carry Select Adder &       & $\surd$ \\
Online Arithmetic Adder  &       & $\surd$ \\
\hline
\hline
\end{tabular}
\end{center}
\vspace{-2ex}
\end{table}
\normalsize

\section{Evaluation of Optimum Adder Structure}\label{sec:optimum_adder}
Generally, our evaluation could be of interest to a circuit designer in several ways. For example, suppose the algorithm designer wish the circuit to run at a given frequency within a specific area budget, while achieving the minimum output error. Besides, for many applications, the circuit designer would wish the circuit to operate as fast as possible with the minimum resource usage, whilst a certain error budget can be tolerated. In both cases, decisions must be made on which adder structure achieves this minimum, and which design scenario should be adopted.
%whether the circuit should be overclocked.

\subsection{Optimum Adder with Given Frequency and Area Requirements}\label{subsec:adder_freq_area}
